sWGS QC

### 2019-08-06 sWGS QC.Rmd
...

library(pander)
library(dplyr)
library(ggplot2)
library(gridExtra)

dir1 = '~/Data/Ellie/Analysis/multipcf_plots_fitted_perPatient'
dir2 = '~/Data/Ellie/Analysis/val_test/'

@@

```{r, echo=F}
res.variance = read.table(paste(dir1,'resid_var.txt',sep='/'),header=T)
#pander(head(res.variance[,grep('var1sd',colnames(res.variance))]), caption='Residuals across segments per sample')
pander(head(res.variance[,grep('varMAD',colnames(res.variance))]), caption='Residuals across segments per sample')
```

@@

```{r, echo=F}
pander( arrange(subset(res.variance, round(varMAD_median,3) > cutoff, select=c(pid,sample, varMAD_median,varMAD_sd, varMAD_Q1,varMAD_Q3)), pid, varMAD_median) )

failedQC = arrange(subset(res.variance, round(varMAD_median,3) > cutoff, select=c(pid,sample, varMAD_median,varMAD_sd, varMAD_Q1,varMAD_Q3)), pid, varMAD_median)

@@

Cutoff from above dataset applied to the validation set.

```{r, echo=F}
list.files(dir2

resVal = read.table(paste(dir2,'resid_var.txt',sep='/'),header=T)
pander(head(resVal[,grep('var1',colnames(resVal))]), caption='Residuals across segments per sample')

resVal = arrange(resVal,varMAD_median)

@@

failedQCVal = arrange(subset(resVal, round(varMAD_median,3) > cutoff, select=c(pid,sample, varMAD_median,varMAD_sd, varMAD_Q1,varMAD_Q3)), pid, varMAD_median)

setdiff(as.character(resVal$sample),as.character(failedQCVal$sample))

@@

```{r}
res.variance[c('pid','sample')] = lapply(res.variance[c('pid','sample')], as.character)
resVal[c('pid','sample')] = lapply(resVal[c('pid','sample')], as.character)

@@

ar = rbind(res.variance, resVal)
ar$cohort = factor(ar$cohort)

@@

fc = arrange(subset(ar, round(varMAD_median,3) > newcutoff, select=c(pid,sample, varMAD_median,varMAD_sd, varMAD_Q1,varMAD_Q3)), pid, varMAD_median)

>>>>

library(tidyverse)
library(gridExtra)
library(kableExtra)

dir1 = '~/Data/Ellie/Analysis/multipcf_perPatient/'
dir2 = '~/Data/Ellie/Analysis/validation//'

@@

```{r, echo=F}

res.variance = do.call(bind_rows, lapply(list.files(dir1, 'residuals', recursive = T, full.names = T), function(f) {
  read_tsv(f, col_types = 'ccdddddl')
}))

#pander(head(res.variance[,grep('var1sd',colnames(res.variance))]), caption='Residuals across segments per sample')
kable(head(res.variance), caption='Example residuals across segments per sample') %>% kable_styling('striped')
```

@@

```{r, echo=F}
res.variance %>% filter( round(varMAD_median,3) > cutoff ) %>% select(patient,sample, varMAD_median) %>% kable(caption=paste0('Samples variance > ',cutoff)) %>% kable_styling('striped')

failedQC = res.variance %>% filter( round(varMAD_median,3) > cutoff ) 
```

@@

Cutoff from above dataset applied to the initial validation set.

```{r, echo=F}
val.file = '~/Data/BarrettsProgressionRisk/QDNAseq/validation/sWGS_validation_batches.xlsx'
sheets = readxl::excel_sheets(val.file)[1:4]
all.val = do.call(bind_rows, lapply(sheets, function(s) {
  readxl::read_xlsx(val.file, s) %>% select(`Hospital Research ID`, matches('Status'), `Block ID`,`Sample Type`, `SLX-ID`, `Index Sequence`, Cohort, Batch, RA, matches('Collection')) %>% dplyr::filter(!is.na(`SLX-ID`) & RA == 'C.Kosmidou') %>% mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) 
}))
all.val = all.val %>% select(-Collection)

pastefun<-function(x) {
  if ( !grepl('SLX-', x) ) x = paste0('SLX-',x)
  return(x)
}

all.val = all.val %>% rowwise %>% mutate_at(vars(`SLX-ID`), list(pastefun) ) %>% ungroup
all.val = all.val %>% mutate(
  `Hospital Research ID` = str_replace_all( str_remove_all(`Hospital Research ID`, " "), '/', '_'), 
  `Index Sequence` = str_replace_all(`Index Sequence`, 'tp', ''),
  Sample = paste(`SLX-ID`,`Index Sequence`,sep='.')
  )

resVal = do.call(bind_rows, lapply(list.files(dir2, 'residuals', recursive = T, full.names = T), function(f) {
  readr::read_tsv(f, col_types = 'cdddddl', col_names = F) %>% set_names('sample', 'varMAD_median', 'varMAD_sd', 'varMAD_Q1', 'varMAD_Q3', 'n.segs', 'Pass') %>%  
    mutate(`Patient ID` = as.character(basename(dirname(f))))
})) %>% select(`Patient ID`, sample, matches('varMad'), n.segs, Pass)

resVal = resVal %>% filter(sample %in% all.val$Sample)

resVal %>% kable(caption='Residuals across segments per sample') %>% kable_styling('striped')
resVal = arrange(resVal,varMAD_median)

@@

failedQCVal = resVal %>% filter(varMAD_median > cutoff) %>% arrange(`Patient ID`, varMAD_median)

#setdiff(as.character(resVal$sample),as.character(failedQCVal$sample))

@@

```{r}
res.variance = res.variance %>% mutate_at(vars(patient, sample), as.character)
resVal = resVal %>% dplyr::rename( patient = `Patient ID` ) %>% mutate_at(vars(patient, sample), as.character)

@@

ar = bind_rows( res.variance, resVal ) %>% mutate(cohort = factor(cohort))

@@

fc = ar %>% filter(round(varMAD_median,3) > newcutoff) %>% select(patient, sample, varMAD_median,varMAD_sd, varMAD_Q1,varMAD_Q3) %>% arrange(patient, varMAD_median)
```
